### 2019-06-27 manuscript/07-supplement.Rmd
Tried to make correct output formats

# Supplementary Information

## p50 and p80

We decided to include values of P50 and P80 in the leaf traits model, defined by [@anderegg_meta-analysis_2016] as the water potentials at which a species loses 50% and 88% [80% by proxy], respectively, of hydraulic conductivity. Values were calculated by (**insert new methods here??**), and were only available for six species (*C. glabra*, *L. tulipifera*, *Q. alba*, *Q. prinus*, *Q. rubra*, and *Q. velutina*). Because of this, the model runs were considered to be incomplete due to the exclusion of the other 8 species. Results revealed neither p50 nor p80 to be significant, thus for the full analysis we decided to drop the two traits in order to include all species in the full analysis.

(see Issue #32)

**other sources???**