### 2019-06-29 manuscript/index.Rmd
changed author order

---
title: "Tree size, exposure, and hydraulic traits interactively shape drought response in a temperate broadleaf forest"
author: "Ian McGregor, Ryan Helcoski, Norbert Kunert, Alan Tepley, Valentine Herrmann, Joseph Zailaa, Neil Pederson, Lawren Sack, Krista Anderson-Teixeira"
output:
  html_document:
    fig.width: 12
    fig.height: 8
  gitbook:
    fig_caption: true
documentclass: book
site: bookdown::bookdown_site
bibliography: [book.bib, packages.bib]
biblio-style: apalike
link-citations: yes
github-repo: manuscript
description: "Article for Ian's research during his time in Anderson-Teixeira lab"
---

```{r include=FALSE}
# automatically create a bib database for R packages
knitr::write_bib(c(
  .packages(), 'bookdown', 'knitr', 'rmarkdown', 'lme4', 'AICcmodavg', 'car', 'piecewiseSEM', 'MuMIn', 'stringr', 'dplyr', 'ggplot2', 'devtools', 'rgdal', 'broom', 'sf', 'ggthemes', 'rgeos', 'RCurl', 'readxl', 'pointRes', 'dplR', 'data.table', 'tools', 'reshape2', 'kableExtra'
), 'packages.bib')
```

<!-- this code makes the .md file but not a correctly-formatted html -->
<!-- output:  -->
<!--   bookdown::html_book: -->
<!--   bookdown::gitbook: -->

<!-- this code makes a correctly-formatted html file but not a .md file -->
<!-- output:  -->
<!--  html_document: -->
<!--     theme: !expr sample(c("yeti", "united", "lumen"), 1) -->
<!--  gitbook: -->
<!--     highlight: tango -->