### 2019-07-01 manuscript/02-intro.Rmd
Update 02-intro.Rmd

# Introduction

Understanding how and why trees respond to drought is critical to predicting forest drought responses and climate change feedbacks.

Forests are diverse in terms of tree sizes and functional traits, and it is known that trees varying in size and functional traits respond differently to drought (e.g., [@bennett_larger_2015]; REFS). Therefore, in order to understand whole-forest response to drought, we need to know how responses vary by tree size/ species. To do so, there are three fundamental questions that must be addressed:

*First, what drives the observed tendency for large trees to suffer more during drought?*  
[@bennett_larger_2015] showed that in forests globally, large trees suffer greater growth reductions during drought. However, this analysis quantified tree size based on DBH, which has no direct mechanistic meaning. This study proposed two major mechanisms (besides insects): (1) inherently greater biophysical challenge of being tall; (2) greater exposure of the crowns of large trees. It is also expected that roots play a role. These hypotheses need to be tested.

*Second, how do species' traits influence drought response?* 
Analyzed drought responses on the species level doesn't get at mechanisms and is not feasible in diverse forests. The solution is a trait-based approach. Leaf hydraulic traits hold more promise than more commonly/ traditionally measured traits such as wood density and SLA (Medeiros et al.).

*Third, how do tree size and functional traits interact to influence drought response?*
Its possible that pattern observed by Bennett et al. could be caused by smaller trees being more drought resistant. Alternatively, larger trees may have more drought-resistant traits as adpatations to greater biophysical challenges.

*Questions & Hypotheses*

**1. Why do larger trees suffer greater growth declines during drought?**

1.0. Our forest displays the same trend as most forests globally [@bennett_larger_2015]. (Note that Bennett et al. 2015 identified only one study on tree growth responses to drought in the Eastern US temperate deciduous biome. We know little about how tree size shapes drought response in this biome.)

1.1. Height is a strong predictor of drought stress.

    * P1.1-Drought response increases with height at time of drought.

1.2. Large trees suffer more during drought because of greater exposure (to radiation, wind, etc.)--either in relation to neighboring trees or because of position on landscape.

    * P1.2a- Trees currently in a dominant canopy position suffered more during drought.

1.3. Rooting volume/depth relative to water sources are critical in drought response. Effects of drought on larger trees are mediated by the fact that large trees have better access to water.

     * P1.3a- drought response increases with elevation and/or distance from stream
     * P1.3b- There is a dbh\*elev interaction, elevation (relative to stream?) matters less for big trees with larger (and potentially deeper) root system.

**2.Do species functional traits predict drought response?**

     * P2.1- TLP correlates negatively with drought resistance (NEVER tested), 
     * P2.2- diffuse porous species are more sensitive than ring porous (previously observed in eastern dec forests- Elliot et al. 2015, Friedrichs et al. 2009)

**3.How are functional traits distributed across size classes, and how does this affect size-resistance relationship?**

3.1. Larger/ more exposed trees have more drought resistant traits, meaning size effects are buffered by traits  
3.1_alt. Larger trees suffer more because they have more drought vulnerable traits.

     * P3.1a- TLP is lower (larger negative) in taller/canopy trees
     * P3.1b- diffuse porous species more common in understory

     * P3.2- Inclusion of TLP / rp in model does not eliminate or significantly reduce effect of tree size.