### 2019-07-30 manuscript/01-abstract.Rmd
Update 01-abstract.Rmd
pasted in relevant content from AGU abstract

# Abstract

*content from AGU abstract:*

Here, we integrate forest census data from a 25.6-ha ForestGEO plot in Virginia (USA), tree-ring records for 12 species representing ##% of woody productivity, leaf hydraulic trait measurements, and microhabitat data.

Individual-level growth responses to three droughts (1964-66, 1977, 1999) were stronger among taller trees in dominant canopy positions, those in wetter microsites, and for more drought-sensitive species as assessed by leaf traits (turgor loss at less negative leaf water potential, greater shrinkage with leaf dehydration), with substantial variation in the best predictor variables across given droughts.

We conclude that when droughts occur, large dominant trees, drought sensitive species, and individuals in moist microhabitats are likely to be most strongly affected.