### 2019-10-12 test.Rmd
Made sample for question on rstudio community

---
title: "Test"
output: pdf_document
---

``` {r} 
dat <- data.frame(variable = c(rep("turgor loss point", 10), rep("volume", 10)),
                    unit = c(rep("$\\pi_{tlp}$", 10), rep("$cm^3$", 10)),
                    value = c(round(runif(10), 2), round(runif(10,20,30),2)))

kableExtra::kable(dat, booktabs = TRUE, format = 'latex', escape=FALSE)
```
Printing in the table leaves the raw symbology, but if I want to discuss the use of the volume ($cm^3$) and the turgor loss point ($\pi_{tlp}$) in the text itself, the symbols render without a problem.