### 2020-02-20 manuscript/test.Rmd

---
title: "test"
output: pdf_document
header-includes: 
- \usepackage{float}
- \usepackage{booktabs}
- \usepackage{pdflscape}
- \newcommand{\blandscape}{\begin{landscape}}
- \newcommand{\elandscape}{\end{landscape}}
- \usepackage{caption}
- \captionsetup[table]{font=small}
- \captionsetup[figure]{font=small}
- \captionsetup[table]{labelformat=empty}
- \captionsetup[figure]{labelformat=empty}
- \usepackage{dcolumn} #for sidewaystable
---

```{r, include=FALSE}
options(tinytex.verbose = TRUE)
options(knitr.table.format = "latex")
```

```{r Table 2, eval = TRUE, echo=FALSE, warning=FALSE}
library(knitr)
library(kableExtra)
table2 <- read.csv("tables_figures/table2_variables_table.csv", stringsAsFactors = FALSE, check.names = FALSE)
kable(table2, booktabs = TRUE, caption = "Table 2. Summary of variables.", 
        escape=FALSE) %>%
  kable_styling()
```