### 2020-02-20 manuscript/test.tex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={test},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{float}
\usepackage{booktabs}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

\title{test}
\author{}
\date{\vspace{-2.5em}}

\begin{document}
\maketitle

\begin{table}

\caption{\label{tab:Table 2}Table 2. Summary of variables.}
\centering
\begin{tabular}[t]{llllll}
\toprule
variable & symbol & units & description & category & n\\
\midrule
drought resistance & Rt & - & ratio of growth during drought year to mean growth of the 5 years prior. & - & 1596\\
drought year & Y & - & year of drought & 1966 & 478\\
 &  &  &  & 1977 & 547\\
 &  &  &  & 1999 & 571\\
diameter breast height & DBH & cm & DBH in drought year & - & all\\
\addlinespace
height & H & m & estimated H in drought year & - & all\\
crown position & CP & - & 2018 crown position & dominant (D) & 31\\
 &  &  &  & co-dominant (C) & 231\\
 &  &  &  & intermediate (I) & 224\\
 &  &  &  & suppressed (S) & 101\\
\addlinespace
topographic wetness index & TWI & - & steady-state wetness index based on slope and upstream contributing area & - & all\\
wood density & WD & g cm-3 & dry mass of a unit volume of fresh wood & - & all\\
leaf mass per area & LMA & kg m-2 & ratio of leaf dry mass to fresh leaf area & - & all\\
xylem porosity & XP & - & vessel arrangement in xylem & ring (R) & 408\\
 &  &  &  & semi-ring (SR) & 31\\
\addlinespace
 &  &  &  & diffuse (D) & 178\\
turgor loss point & $\psi_{tlp}$ & MPa & water potential at which leaves wilt & - & all\\
percent loss area & PLA & % & percent loss of leaf area upon dessication & - & all\\
\bottomrule
\end{tabular}
\end{table}

\end{document}