### 2020-03-21 non_publication/manuscript/original_bookdown_files/archive_methods.Rmd
moved to non_publication folder (iii)
Files that are not being used in the final manuscript were moved into a non_publication folder as a holding step. Next step is dependent on each file, either delete, keep here or consolidate further.

---
title: "archive_methods"
output: html_document
---
This document is meant to keep track of methods, even though we're not necessarily putting them in the paper.

## Sapwood ratio
Sapwood ratio also used the bark thickness data, and was also calculated for 2008, where
```
sapwood_area = total wood area (without bark) - heartwood area, and
sapwood_Ratio = sapwood area / total wood area (without bark)
```
Once sapwood area for each tree was calculated, log-log regression equations were determined on a per-species basis, which were then used